---
title: "ESM 244 | Homework 2"
author: "Hanna Buechi"
date: "2/12/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r data_packages, warning=FALSE, error=FALSE, message=FALSE}

library(tidyverse)
library(ggrepel)
library(lubridate)
library(RColorBrewer)

whales <- read_csv("captive_cetacean.csv")
fish <- read_csv("fish_catch.csv")

```

###Task 1. Data wrangling and visualization - cetaceans in captivity


####Initial data wrangling
```{r whales_wrangling, warning=FALSE, error=FALSE, message=FALSE, results='hide'}

summary(whales) # everything is character/string data
# interesting time data is birth year (single data point) and transfers (could tell a story about individuals but hard to parse)
# something more than a single histogram... histogram with some other elements...

whales$originDate <- as.Date(whales$originDate, "%m/%d/%y") # lower case "y" keeps date as 19xx =)
whales$originYear <- lubridate::year(whales$originDate)

# Check to make sure dates expanded correctly

as.data.frame(table(whales$originYear))

# NOPE --> wrong years: 2049, 2052, 2053, 2055, 2056, 2059, 2060, 2061, 2062, 2063, 2064, 2065, 2066, 2067, 2068

whales <- whales %>%
  mutate(originYear=replace(originYear, originYear==2049, 1949)) %>%  
  mutate(originYear=replace(originYear, originYear==2052, 1952)) %>%
  mutate(originYear=replace(originYear, originYear==2053, 1953)) %>%
  mutate(originYear=replace(originYear, originYear==2055, 1955)) %>%
  mutate(originYear=replace(originYear, originYear==2056, 1956)) %>%
  mutate(originYear=replace(originYear, originYear==2058, 1958)) %>%
  mutate(originYear=replace(originYear, originYear==2059, 1959)) %>%
  mutate(originYear=replace(originYear, originYear==2060, 1960)) %>%
  mutate(originYear=replace(originYear, originYear==2061, 1961)) %>%
  mutate(originYear=replace(originYear, originYear==2062, 1962)) %>%
  mutate(originYear=replace(originYear, originYear==2063, 1963)) %>%
  mutate(originYear=replace(originYear, originYear==2064, 1964)) %>%
  mutate(originYear=replace(originYear, originYear==2065, 1965)) %>%
  mutate(originYear=replace(originYear, originYear==2066, 1966)) %>%
  mutate(originYear=replace(originYear, originYear==2067, 1967)) %>%
  mutate(originYear=replace(originYear, originYear==2068, 1968))

# df of cetaceans that were captured

capture <- whales %>% 
  filter(acquisition == "Capture")

# df of cetaceans that were captured in 1972

capture72 <- filter(capture, originYear == 1972)

```


####Group species by "type"
#####*I didn't end up using this classification*
```{r species, warning=FALSE, error=FALSE, message=FALSE}

# Group by dolphins, orcas, belugas, other whales
# Dolphin: "Bottlenose" "Commerson's" "Common" "Pacific White-Sided" "Pantropical Spotted" "Risso's" "River, Amazon" "Rough-Toothed" "Spinner" "Spotted, Atlantic" "Tucuxi" "White-Beaked" "White-sided, Pacific"
# Orca: "Killer Whale; Orca" 
# Beluga whale: "Beluga" and "Beluga whale"
# Other species: "False Killer Whale" "Melon-Headed Whale" "Northern Right Whale" "Pilot, Short-fin" "Pilot, Short-finned" "Pseudorca" "Pygmy Killer Whale" "Short-Finned Pilot Whale" "Unspecified Pilot Whales"

capture <- capture %>% 
  mutate(Type = ifelse(species == "Bottlenose", "Dolphins",
                ifelse(species == "Commerson's", "Dolphins",
                ifelse(species == "Common", "Dolphins",
                ifelse(species == "Pacific White-Sided", "Dolphins",
                ifelse(species == "Pantropical Spotted", "Dolphins",
                ifelse(species == "Risso's", "Dolphins",
                ifelse(species == "River, Amazon", "Dolphins",
                ifelse(species == "Rough-Toothed", "Dolphins",
                ifelse(species == "Spinner", "Dolphins",
                ifelse(species == "Rough-Toothed", "Dolphins",
                ifelse(species == "Spotted, Atlantic", "Dolphins",
                ifelse(species == "Tucuxi", "Dolphins",
                ifelse(species == "White-Beaked", "Dolphins",
                ifelse(species == "White-sided, Pacific", "Dolphins",
                ifelse(species == "Killer Whale; Orca", "Orcas",
                ifelse(species == "Beluga", "Beluga whales",
                ifelse(species == "Beluga whale", "Beluga whales",
                "Others"))))))))))))))))))

# Did not use this in visualization

```


####Visualize!
```{r capture_plot, warning=FALSE, error=FALSE, message=FALSE}

# The Marine Mammal Protection Act was effective December 21, 1972 -- introduced December 4, 1971
# NOAA issues permits for importing marine mammals and collecting them from the wild. No permit needed to transfer between US facilities (except approval from APHIS under Animal Welfare Act).

# check what is going on in 1972

 # 165 animals

ggplot(capture, aes(x = originYear)) +
  geom_bar() +
  labs(title="Regulating Cetaceans in Captivity", 
       subtitle="On December 21, 1972, The Marine Mammal Protection Act went into effect. It requires a permit\nto capture a cetacean for scientific or display purposes.\n") +
  theme(plot.title = element_text(face = "bold")) +
  xlab("Year") +
  ylab("Cetaceans captured") +
  geom_segment(aes(x = 1950, y = 170, xend = 1972.5, yend = 170)) +
  geom_segment(aes(x = 1973, y = 170, xend = 2000, yend = 170)) +
  geom_text(x=1960, y=160, label="Before MMPA") +
  geom_text(x=1985, y=160, label="After MMPA") +
  geom_text(x=1960, y=100, label="165 whales and dolphins\nwere captured from the wild\nin 1972.") +
  annotate("rect", xmin = 1951, xmax = 1969, ymin = 80, ymax = 120,
  alpha = .2, fill = "blue") +
  theme_classic()


```

###Task 2. Parameter Estimation - Wild Fish Catch

Data source: Global wild fish catch and aquaculture production, compiled by Earth Policy Institute
with 1950-2010 from U.N. Food and Agriculture Organization (FAO), Global Capture Production
and Global Aquaculture Production, electronic databases, at
www.fao.org/fishery/topic/16140/en. Catch is in millions of tons.

```{r fish_wrangling}

# set 1950 as year 0

# make a vector with 63 numbers, starting at 0, column bind to fish df

time <- c(0:62)

fish <- fish %>%
  slice(1:63) %>% # remove text in the last few rows that I missed when cleaning the csv in Excel
  cbind(time)

```
####Part a. Exploratory graph

```{r wild_time}

# wild catch over time

ggplot(fish, aes(x = Year, y = wild)) +
  geom_point() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
  theme_minimal()

```





